---
title: "Agri-food CO2 Emission"
author: "Eva Marudur"
date: "2024-06-14"
output: html_document
theme: cosmo
    highlight: breezedark
    toc: true
    toc_float:
      collapsed: false
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(dplyr) # data prep
library(lubridate) # date data prep
library(ggplot2) # visualisasi statis
library(plotly) # plot interaktif
library(glue) # setting tooltip
library(scales) # mengatur skala pada plot
```
## Membaca data 

```{r}
# Please type your code here
CO_Emission <- read.csv("data_input/Agrofood_co2_emission.csv")
CO_Emission
```

#Inspeksi Data Tipe dari CO_Emission:


```{r}
CO_Emission %>%
  glimpse()
```


vids_clean_quiz <- 
vids_quiz %>%
  mutate_at(.vars =  c("channel_title", "category_id"),
            .funs = as.factor)%>% 
mutate(year = ymd(year),
       publish_time = ymd_hms(publish_time))



```{r}
CO_Emission_Clean <- 
CO_Emission %>%
  mutate_at(.vars =  c("Area", "Region"),
            .funs = as.factor)

```

  
```{r}
CO_Emission_Clean %>%
  glimpse()
```
  

# -------- TAB 1

## plotly: CO_Emissison per area per year


```{r}
#library(dplyr)
CO_Emission_Region <- 
CO_Emission_Clean %>%
  group_by(Region,Year) %>% 
  summarise(CO_Emmision_Total = sum(total_emission)) %>% 
  arrange(desc(CO_Emmision_Total))%>% 
  mutate(text = paste0("CO_Emmision_Total: ", CO_Emmision_Total, " kt"))
```
#visualisasi Static

```{r}
CO_Emission_Region_plot <- 
  ggplot(CO_Emission_Region,mapping =  aes(x=reorder(Region, CO_Emmision_Total), 
                               y= CO_Emmision_Total,
                              fill = CO_Emmision_Total,
                              text = text)) +
            geom_col(aes(fill = CO_Emmision_Total), show.legend = F) +
            coord_flip()+
            labs(title = "Total CO Emission per Region",
                  x = "Total CO Emission",
                  y = "") + 
            scale_fill_gradient(low = "#F78181",
                      high = "#3B0B0B") +
theme_minimal() +
theme(legend.position = "none")
```



```{r}
ggplotly(CO_Emission_Region_plot, tooltip = 'text')
```


```{r}
CO_Emission_Region_plot_interactive <- ggplotly(CO_Emission_Region_plot, tooltip = 'text')
```


```{r}
CO_Emission_Region_plot_interactive
```


CO_Emission_Area_plot <- ggplot(ef_region, aes(x=reorder(Region, Ecological.Footprint), y=Ecological.Footprint, text = text)) +
            geom_col(aes(fill=Ecological.Footprint), show.legend = F) +
            coord_flip() +
            labs(title = NULL,
                 y = "global hectares (gha)",
                 x = NULL) +
            scale_y_continuous(limits = c(0, 150),
                           breaks = seq(0,150, 25)) +
            scale_fill_gradient(low = "#F78181", high = "#3B0B0B") +
            theme(plot.title = element_text(face = "bold", size = 14, hjust = 0.04),
              axis.ticks.y = element_blank(),
              panel.background = element_rect(fill = "#ffffff"), 
              panel.grid.major.x = element_line(colour = "grey"),
              axis.line.x = element_line(color = "grey"),
              axis.text = ele
